---
title: "p8105_hw2_hx2263"
author: "Tiffany Xi"
date: 2018-09-27
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

#Problem 1

Import and clean NYC transit csv files

```{r df_import_clean, message = FALSE}
nyc_transit = 
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, starts_with("route"), entry,          vending, entrance_type, ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```

*Description:*

This nyc_transit dataset contains variables named `r colnames(nyc_transit)`. 

My data cleaning steps are: first clean the column names into snake case with function janitor::clean_names; then use 'select' function to select the variables I am interested in. Subsequently, use recode to convert the character variable 'YES' and 'NO' in 'entry' column into logical variable 'TRUE' and 'FALSE'.

The dimension of the resulting dataset is `r nrow(nyc_transit)` * `r ncol(nyc_transit)`. I think these data are untidy, we need to put all routes into one column.

* Answers:

1. There are `r nrow(distinct(nyc_transit, line, station_name))` distinct stations.

2. The number of disctinct stations which are ADA compliant is `r nrow(nyc_transit %>% distinct(line, station_name, .keep_all = TRUE) %>% filter(ada == "TRUE"))`. From the original dataset, the number of stations which are ADA compliant is `r nrow(nyc_transit %>% filter(ada == "TRUE"))`.
 
3. The proportion of station entrances / exits without vending allow entrance is `r nrow(nyc_transit %>% filter(vending == 'NO' & entry == TRUE))/nrow(nyc_transit %>% filter(vending == 'NO'))`.

* Reformat data

*1.* route number and route name are distinct variables
```{r reformat_1}
nyc_transit_tidy = 
  nyc_transit %>% 
  gather(key = "route_number", value = "route_name", route1:route11) %>% 
  separate(route_number, into = c("route_str", "route_num"), sep = 5) %>% 
  select(-route_str)
nyc_transit_tidy
```

*2.* `r nrow(nyc_transit_tidy %>% distinct(line, station_name, .keep_all = TRUE) %>% filter(route_name == "A"))` distinct stations serve the A train.

*3.* Of the stations that serve the A train, `r nrow(nyc_transit_tidy %>% distinct(line, station_name, .keep_all = TRUE) %>% filter(route_name == "A") %>% filter(ada == "TRUE"))` of them are ADA compliant.


#Problem 2

1.Read and clean the Mr. Trash Wheel sheet.

```{r trashwheel_import}
trash_wheel = 
  readxl::read_excel(path = "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                     sheet = "Mr. Trash Wheel", 
                     range = "A2:N338") %>%
  janitor::clean_names() %>% 
  filter(!is.na(dumpster)) %>% 
  mutate(sports_balls = round(sports_balls, digits = 0)) %>% 
  mutate(sports_balls = as.integer(sports_balls))
trash_wheel
```
  
2.Read and clean precipitation data for 2016 and 2017

```{r data_16_17}
prcp_2016 = 
  readxl::read_excel(path = "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                     sheet = "2016 Precipitation", 
                     range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(month)) %>% 
  filter(!is.na(total)) %>% 
  mutate(year = 2016) 

prcp_2017 = 
  readxl::read_excel(path = "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                     sheet = "2017 Precipitation", 
                     range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(month)) %>% 
  filter(!is.na(total)) %>% 
  mutate(year = 2017) 
```

3.Combine datasets and convert month to a character variable

```{r comb_convert}
prcp_2016 = mutate(prcp_2016, month = month.name)
prcp_2017 = mutate(prcp_2017, month = month.name)
combo_16_17 = full_join(prcp_2016, prcp_2017, by = c("month", "total", "year"))
combo_16_17
```

* Description:

In the dataset `trash_wheel`, the number of observations is `r nrow(trash_wheel)`, and the number of variables is `r ncol(trash_wheel)`.

* Key variable:

1.The mean of the number of glass bottles in 2014, 2015, 2016 are `r mean(trash_wheel$glass_bottles[which(trash_wheel$year == 2014)])`, `r round(mean(trash_wheel$glass_bottles[which(trash_wheel$year == 2015)]), digits = 2)` and `r round(mean(trash_wheel$glass_bottles[which(trash_wheel$year == 2016)]), digits = 2)`.

2.The median of the number of grocery bags in 2015, 2016, 2017 are `r median(trash_wheel$grocery_bags[which(trash_wheel$year == 2015)])`, `r median(trash_wheel$grocery_bags[which(trash_wheel$year == 2016)])` and `r median(trash_wheel$grocery_bags[which(trash_wheel$year == 2017)])`.

3.The maximum of the number of cigarette butts in the year 2015 is `r max(trash_wheel$cigarette_butts[which(trash_wheel$year == 2015)])`.


In the dataset `combo_16_17`, the number of observations is `r nrow(combo_16_17)`, and the number of variables is `r ncol(combo_16_17)`.

* Key variable:

1.The mean of precipitation data in year 2016 is `r round(mean(combo_16_17$total[which(combo_16_17$year == 2016)]), digits = 2)`.

2.The median of precipitation data in year 2017 is `r median(combo_16_17$total[which(combo_16_17$year == 2017)])`.

3.The standard deviation of precipitation data in year 2016 is `r round(sd(combo_16_17$total[which(combo_16_17$year == 2016)]), digits = 2)`.

* For available data, the total precipitation in 2017 was `r sum(prcp_2017$total[prcp_2017$year == "2017"])` inches.

* The median number of sports balls in a dumpster in 2016 was `r median((trash_wheel %>% filter(trash_wheel$year == 2016))$sports_balls)`.


#Problem 3

Load the data from the p8105.datasets package

```{r data_import, message = FALSE}
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)
```

Clean the data and reformat according to requirement

```{r brfss_clean}
brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  select(-class, -topic, -question, -sample_size, -(confidence_limit_low:geo_location)) %>%
  spread(key = response, value = data_value) %>%
  janitor::clean_names() %>% 
  select("year", "locationabbr", "locationdesc", "excellent", "very_good", "good", "fair", "poor") %>% 
  mutate(prop_ex_vg = (excellent + very_good))

```


* Answers:

What state is observed the most?

```{r most_observed}
which.max(table(brfss$locationabbr))
```

* `r nrow(brfss %>% distinct(locationdesc, .keep_all = TRUE))` unique locations are included in the dataset. The number of states are `r nrow(brfss%>%distinct(locationabbr, .keep_all = TRUE))`, so, yes, every state is represented. NJ(New Jersey) is observed the most.

* In 2002, the median of the “Excellent” response value is `r median((brfss %>% filter(year == 2002))$excellent, na.rm = TRUE)`.
                                          
* Histogram of “Excellent” response values in the year 2002

```{r, warning = FALSE, message = FALSE}
brfss %>% 
  filter(year == 2002) %>% 
  ggplot(aes(x = excellent)) + 
  geom_histogram() +
  labs(
    title = "Histogram of Excellent response values in 2002",
    x = "Excellent response values",
    y = "Counts",
    caption = "Data from the p8105.datasets package"
  ) +
  scale_x_continuous(breaks = c(15, 20, 25, 30, 35, 40), 
                     labels = c("15", "20", "25", "30", "35", "40")) +
  theme_bw()
```

* Scatterplot of the proportion of “Excellent” response values in New York County and Queens County (both in NY State) in each year from 2002 to 2010

```{r}
  brfss %>% 
  filter(locationabbr == "NY") %>% 
  filter(locationdesc == "NY - New York County" | locationdesc == "NY - Queens County") %>% 
  ggplot(aes(x = year, y = excellent)) + 
  geom_point(aes(color = locationdesc)) +
  labs(
    title = "Scatterplot of the proportion of Excellent ",
    x = "Year",
    y = "The Proportion of Excellent",
    caption = "Data from the p8105.datasets package"
  ) +
  scale_x_continuous(breaks = c(2002, 2003, 2004, 2005, 2006, 
                                2007, 2008, 2009, 2010), 
                     labels = c("2002", "2003", "2004", "2005", "2006", 
                                "2007", "2008", "2009", "2010")) +
  theme_bw() +
  theme(legend.position = "bottom")
```


